Drought vulnerability of tree growth

### 2024-07-10 scripts/explore_analysis.Rmd
updated resistance values plot @teixeirak

resistance_top4 <- top4_noouts %>%
  filter(Cno %in% c(5, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(
    yr=ifelse(Cno==5, 2010, 2015),
        resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-avg_inc,
         resistance.prop = (inc_annual-avg_inc)/avg_inc)%>%
         ungroup()%>%
  filter(!is.infinite(resistance.div))%>%
         group_by(Cno)%>%
         #making quantiles for each year
         dplyr::mutate(
         quantiles_div=cut(resistance.div,
         breaks=quantile(resistance.div, 
         #probs=c(0.25, 0.5, 0.75, 1), 
         probs=seq(0, 1, 0.25),
         na.rm=T),
         labels=c("1", "2", "3", "4")))

@@

#2010 drought--------

resistance_2010 <- dendro_inc %>%
filter(Cno %in% c(3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2010<-merge(resistance_2010, select(dec_williams, -"X"), by="sp", all.x=T)

resistance_2010_div<-ggplot(data = resistance_2010, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  geom_vline(xintercept = 0, lty=2)+
  theme_bw()

resistance_2010_dif<-ggplot(data = resistance_2010, aes(x=resistance.dif)) +
  geom_density(alpha=0.5) +
  theme_bw()

#2015 drought---------

resistance_2015 <- dendro_2015 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2015<-merge(resistance_2015, select(dec_williams, -"X"), by="sp", all.x=T)

resistance_2015_div<-ggplot(data = resistance_2015, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  geom_vline(xintercept = 0, lty=2)+
  theme_bw()

resistance_2015_dif<-ggplot(data = resistance_2015, aes(x=resistance.dif)) +
  geom_density(alpha=0.5) +
  theme_bw()

#2019 drought--------------

resistance_2019 <- dendro_2019 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/inc_min1,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/inc_min1)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2019<-merge(resistance_2019, select(dec_williams, -"X"), by="sp", all.x=T)

resistance_2019_div<-ggplot(data = resistance_2019, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  theme_bw()

resistance_2019_dif<-ggplot(data = resistance_2019, aes(x=resistance.dif)) +
  geom_density(alpha=0.5) +
  theme_bw()

png("../doc/display/resistance_dif.png", width=4, height=8, units="in", res=300)
grid.arrange(resistance_2010_dif, resistance_2015_dif, resistance_2019_dif, nrow=3)
dev.off()

library(gridExtra)

#png("growth-precip-thailand/doc/display/resistance_div.png", width=4, height=8, units="in", res=300)
#grid.arrange(resistance_2010_div, resistance_2015_div, resistance_2019_div, nrow=3)
#dev.off()

png("growth-precip-thailand/doc/display/resistance_div.png", width=8, height=4, units="in", res=300)
grid.arrange(resistance_2010_div, resistance_2015_div, nrow=1)
dev.off()

@@

quantiles_2010_dif<-quantile(resistance_2010$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2015_dif<-quantile(resistance_2015$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

quantiles_2019_dif<-quantile(resistance_2019$resistance.dif, 
                         probs=c(0.25, 0.5, 0.75, 1))

resistance_2010$quantile<-ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[1], 1, ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[2], 2,
            ifelse(resistance_2010$resistance.dif<quantiles_2010_dif[3], 3, 4))

                                 )

resistance_2015$quantile<-ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[1], 1, ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[2], 2,
            ifelse(resistance_2015$resistance.dif<quantiles_2015_dif[3], 3, 4))

                                 )

resistance_2019$quantile<-ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[1], 1, ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[2], 2,
            ifelse(resistance_2019$resistance.dif<quantiles_2019_dif[3], 3, 4))

                                 )

quantiles_2010_div<-quantile(resistance_2010$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

quantiles_2015_div<-quantile(resistance_2015$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

quantiles_2019_div<-quantile(resistance_2019$resistance.div, 
                         probs=c(0.25, 0.5, 0.75, 1), na.rm=T)

resistance_2010$quantile.div<-ifelse(resistance_2010$resistance.div<quantiles_2010_div[1], 1, ifelse(resistance_2010$resistance.div<quantiles_2010_div[2], 2,
            ifelse(resistance_2010$resistance.div<quantiles_2010_div[3], 3, 4))

                                 )

resistance_2015$quantile.div<-ifelse(resistance_2015$resistance.div<quantiles_2015_div[1], 1, ifelse(resistance_2015$resistance.div<quantiles_2015_div[2], 2,
            ifelse(resistance_2015$resistance.div<quantiles_2015_div[3], 3, 4))

                                 )

resistance_2019$quantile.div<-ifelse(resistance_2019$resistance.div<quantiles_2019_div[1], 1, ifelse(resistance_2019$resistance.div<quantiles_2019_div[2], 2,
            ifelse(resistance_2019$resistance.div<quantiles_2019_div[3], 3, 4))

                                 )

resistance_all<-rbind(resistance_2010, resistance_2015)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010)), rep(2015, nrow(resistance_2015)))

resistance_all<-merge(resistance_all, select(trees, -"SPCODE.UPDATE"), by="Tag")

>>>>

resistance_top4 <- top4_noouts %>%
  filter(Cno %in% c(5, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(
    yr=ifelse(Cno==5, 2010, 2015),
        resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-avg_inc,
         resistance.prop = (inc_annual-avg_inc)/avg_inc)%>%
         ungroup()%>%
  filter(!is.infinite(resistance.div))%>%
         group_by(Cno)%>%
         #making quantiles for each year
         dplyr::mutate(
         quantiles_div=ntile(resistance.div, 4))

@@

resistance <- dendro_inc_clean %>%
  filter(Cno %in% c(5, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(
    yr=ifelse(Cno==5, 2010, 2015),
        resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-avg_inc,
         resistance.prop = (inc_annual-avg_inc)/avg_inc)%>%
         ungroup()%>%
  filter(!is.infinite(resistance.div))%>%
         group_by(Cno)%>%
         #making quantiles for each year
         dplyr::mutate(
         quantiles_div=ntile(resistance.div, 4),
         quantiles_dif=ntile(resistance.dif, 4),
         quantiles_prop=ntile(resistance.prop, 4))

resistance<-merge(resistance, select(trees, -"SPCODE.UPDATE"), by="Tag", all.x=T)
resistance<-merge(resistance, dec_williams, by="sp", all.x=T)

#distribution and map

resistance_div<-ggplot(data = resistance, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  facet_grid(.~yr)+
  xlim(-5, 5)+
  xlab("inc/long-term increment")+
  ggtitle("resistance values for all trees")+
  geom_vline(xintercept = c(0,1), lty=2)+
  theme_bw()

png("growth-precip-thailand/doc/display/resistance_div.png", width=6, height=4, units="in", res=300)
resistance_div
dev.off()

@@
